% This program is free software: you can redistribute it and/or modify it under the terms of the GNU General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version.
%
% This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License along with this program. If not, see <https://www.gnu.org/licenses/>.

\input{common.tex}

\title{Lecture \#11: logging}
\date{\today}
\author{Presented by Jamal Bouajjaj}
\institute{For University of New Haven's Fall 2023 CSCIxx51 Course}

% From https://tex.stackexchange.com/questions/66519/make-each-frame-not-slide-appear-in-the-pdf-bookmarks-with-beamer
\makeatletter
\apptocmd{\beamer@@frametitle}{\only<1>{\bookmark[page=\the\c@page,level=3]{#1}}}%
% {\message{** patching of \string\beamer@@frametitle succeeded **}}%
% {\message{** patching of \string\beamer@@frametitle failed **}}%
\makeatother

\begin{document}

\maketitle

\begin{frame}{Logging Stuff}
  Let's say you have a functional application. Cool!\pause

  Now what if during run-time an error happended? Wouldn't you like to know what occured?
  Both during debugging and deployment?
\end{frame}

\begin{frame}[containsverbatim]{Naive Logging}
  The following is a simple way to "log" stuff to the terminal. The simple print statement.
\begin{minted}[breaklines=true,frame=single]{python}
a = some_function(5, 2)
print(f"a is {a}")
try:
  do_something_else(a, 50)
except ValueError as e:
  print("Exception here! ", e)
\end{minted}
\end{frame}

\begin{frame}{logging Module}
  While the above could be sufficient, there are a couple of issues/downsides:
  \begin{itemize}
    \item If this is a CLI application, the user application will have print-outs that aren't relavent to what the user needs.
    \item There is no distunction between important errors and just debug logs.
    \item What is printed does not get sent to a file, allowing loss of information with no way of getting it.
    \item If a GUI application, there will be no console. Same as above
    \item Exception printing is limited
  \end{itemize}
\end{frame}

\begin{frame}{Solutions}
  Enough rambling, what is a viable solution?\pause
  \begin{itemize}
    \item Force the user to open a command-line when running the application, and requiring them to copy the output if there is an error\pause
    \item Give up!\pause
    \item Ignore any errors, they aren't important.\pause
    \item Use the \textit{logging} module
  \end{itemize}
\end{frame}

\section{logging}

\begin{frame}{logging}
  logging is a standard module that is designed to handle logging stuff in your application.
\end{frame}

\begin{frame}[containsverbatim]{Logger}
  A Logger is an object that handles logging. There is always a root logger (top level), and you can create a logging handler. Each logger can have it's own "settings", as in levels, handlers, etc.

  Getting a logger with the same name will return the same logger.

  A logger name is hierarchal, so root is the top level, then all other loggers are child loggers. A dot indicates a child of the logger name.
  \begin{minted}[breaklines=true,frame=single]{python}
l = logging.getLogger() # root logger
l = logging.getLogger('sensor')   # a logger with name
l = logging.getLogger('sensor.ser')  # a logger with name, child of sensor
\end{minted}
\end{frame}

\begin{frame}{Log Levels}
  There are several levels of log message withing the module\footnote{\url{https://docs.python.org/3/howto/logging.html}}:
  \begin{itemize}
    \item DEBUG
    \item INFO
    \item WARNING
    \item ERROR
    \item CRITICAL
  \end{itemize}
\end{frame}

\begin{frame}[containsverbatim]{Level Example}
  This is how to log with levels
  \begin{minted}[breaklines=true,frame=single]{python}
l = logging.getLogger() # root logger
l.debug("Message")
l.info("Message")
l.warning("Message")
l.error("Message")
l.critical("Message")
# calls to logging (the module) directly will just log with the root logger
logging.debug("Message")
\end{minted}
\end{frame}

\begin{frame}[containsverbatim]{Exception Logging}
  You can also log exceptions, which will also record the traceback! The level is ERROR.
  \begin{minted}[breaklines=true,frame=single]{python}
l = logging.getLogger() # root logger
try:
  a = 5 / 0
except:
  logging.exception("Error!")
\end{minted}
\end{frame}

\begin{frame}{Log Handlers}
  Once a log is emmited (i.e sent), where does it get sent? That is determined by what handlers exist for the logger.

  The handler just handles what to do with a log.

  Each handler can have a level: anything below the set level is ignored for the handler.

  The handler is applied to a logger on a hierarchal basis. So a handler for the root logger will get called for all child loggers, etc.
\end{frame}

\begin{frame}[containsverbatim]{Log Handlers}
  \begin{minted}[breaklines=true,frame=single]{python}
l = logging.getLogger()
hs = logging.StreamHandler()
hs.setLevel(logging.WARNING)
l.addHandler(hs)
fs = logging.FileHandler("file.log")
hs.setLevel(logging.DEBUG)
l.addHandler(fs)
  \end{minted}
\end{frame}

\begin{frame}[containsverbatim]{Log Format}
  You can change what each log handler formats it's output. You can for example have a log output the function name, time, message, etc.
    \begin{minted}[breaklines=true,frame=single]{python}
l = logging.getLogger()
hs = logging.StreamHandler()
hs.setLevel(logging.WARNING)
f = logging.Formatter("%(asctime)s %(levelname)s:%(filename)s:%(funcName)s:%(lineno)s %(message)s")
hs.setFormatter(f)
l.addHandler(hs)
  \end{minted}
\end{frame}

\begin{frame}[containsverbatim]{Logging Without Setup!}
  logging can have a decent amount of setup for it. If you just want something to just log in the strem, and maybe file, you can have logging setup a lot for you by using the following function:
  \begin{minted}[breaklines=true,frame=single]{python}
logging.basicConfig(level=logging.DEBUG)
logging.basicConfig(level=logging.DEBUG, filename="program.log")
\end{minted}
\end{frame}


\begin{frame}[standout]{End}
  The end
\end{frame}

\end{document}
